\documentclass[12pt,a4paper,headsepline,footsepline,DIV13,BCOR12mm]{scrbook}
\usepackage[english,ngerman]{babel}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{scrpage2}
\pagestyle{scrheadings}

%\usepackage[backend=biber,style=unsrt]{biblatex}

\usepackage{graphicx}

\usepackage{setspace}

\usepackage{tikz}
\usetikzlibrary{arrows, shapes}
\usepackage{pgf-umlsd}
\usepackage{amsmath}

% LTL Symbols
\usepackage{latexsym}
\usepackage{MnSymbol}

\usepackage{wrapfig}

\usepackage{listings}
\usepackage{color}
\definecolor{gray}{rgb}{0.4,0.4,0.4}
\definecolor{darkblue}{rgb}{0.0,0.0,0.6}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}

%for bookmark bar
\usepackage[colorlinks,pdfpagelabels,pdfstartview=FitH,bookmarksopen=false,bookmarksnumbered=false,linkcolor=black,plainpages=false,hypertexnames=false,citecolor=black]{hyperref}

\usepackage{algorithm}
\usepackage{algorithmic}



\begin{document}



%PREFACE
\selectlanguage{ngerman}
\input{titlepage} 



\newpage
\thispagestyle{empty}

\centerline{\bfseries ERKLÄRUNG}

\vspace{5cm}
Hiermit versichere ich, dass ich diese Masterarbeit selbständig verfasst habe.
Ich habe dazu keine anderen als die angegebenen Quellen und Hilfsmittel
verwendet.

\vspace{1cm}
\begin{flushleft}
%select german for formatting the date
\selectlanguage{ngerman}
Augsburg, den 16. November 2012 \hfill Ludwig Nägele %\today
\end{flushleft}

\newpage
\thispagestyle{empty}
\mbox{}








\newpage
\onehalfspacing
\selectlanguage{ngerman}
\input{tex/0_zusammenfassung}
\selectlanguage{english}
\input{tex/0_abstract}

\chapter*{Acknowledgment}
%\thispagestyle{empty}

For my master thesis it was my wish to do research abroad in a foreign country. Professor Wolfgang Reif, supervisor of my thesis, supported me in my plan and managed to give me the unique possibility to visit New Zealand for half an year. Therefore I am deeply grateful to him.

My thesis project was hostet at the University of Auckland, New Zealand, by Professor Bruce MacDonald. I want to thank him for the opportunity to stay and to work at his robotics lab, for all the meetings and discussions we had and for all the support he gave to me.

I also want to thank Felix Kaser, a fellow student of mine, who was with me in New Zealand. During several coffee breaks we frequently discussed ideas for this work and he helped me a lot with his good suggestions.

Chandan Datta, PhD student at the University of Auckland and developer of Robostudio, always offered his help regarding the use of Robostudio or general questions about the healthcare robotics project. He furthermore supported me in getting familiar with the lab and its research topics and with Auckland as a city to live in. Especially in the first time after my arrival in New Zealand I was very grateful for this.

Sincere thanks is also given to Alwin Hoffmann and my supervising tutor Andreas Angerer, both doctoral candidates at the University of Augsburg, Germany, for assisting me by discussions and hints. They did not only inspect and correct the thesis writing regularly, but they also assisted me in finding the thesis topic in the very beginning and contributed to the results of this work decisively by providing ideas and feedback during the research phase.

Last but not least I want to thank my father Rudolf Nägele for all the effort he put in reading and correcting my work. Since this work relates to different mathematical concepts and formalisms such as LTL, model checking, RBDL, etc., he had to spend plenty of hours in front of the computer in order to get familiar with all these concepts and to check this work for plausibility and correctness.



%\newpage
%\thispagestyle{empty}
%\mbox{}
\newpage

%select english as language!
%\selectlanguage{english}
%\pagenumbering{roman}
%\pagenumbering{arabic}

\pdfbookmark[1]{Contents}{toc}
\tableofcontents
\newpage
\listoffigures
\newpage
%\listoftables
\lstlistoflistings
%\pagenumbering{arabic}

\graphicspath{{./images/}}
\DeclareGraphicsExtensions{.jpg}


\input{tex/1_introduction}
\input{tex/2_background}
\input{tex/3_goals}
\input{tex/4_visualformalismsforsafetyconstraints}
\input{tex/5_automatedgenerationofsafetyconstraints}
\input{tex/6_theltlcreatorprototype}
\input{tex/7_testscenarioandevaluation}
\input{tex/8_conclusionandfuturework}






%This work evolved from a master thesis project provided by the Elite Graduate Program Software Engineering. 
%For the received support I would like to thank the people of this program and the three patron universities: University of Augsburg, Technical University Munich and Ludwig-Maximilians-University Munich.

\bibliographystyle{unsrt}
\bibliography{Bib}

\end{document}